---
title: "Model Validation"
execute:
  warning: false
  message: false
format:
  html:
    theme: flatly
  pdf:
    fontsize: 11pt
df-print: kable
fig-width: 6
fig-height: 6
---

# Imports

```{r}
#| code-fold: true
#| output: false

# General
library(tidyverse)

# Visualization
library(ggplot2)

# Miscelaneous
library(here)
library(glue)
```

```{r}
#| echo: false

# Helper function to display a plot in Plotly if the report is in HTML format.
render_plot <- function(p) {
    if (knitr::is_html_output()) {
        plotly::ggplotly(p)
    } else {
        p
    }
}
```

# Data Loading

```{r}
aicc <- readr::read_csv2(here("output", "validation", "aicc.csv"))
rmse <- readr::read_csv2(here("output", "validation", "rmse.csv"))
```

{{< pagebreak >}}

# Evaluation
## AICc 

```{r}
#| code-fold: true

# Create a column called `best_model` with the arima model which minimizes AICc, then count them.
# Also, add the mean AICc among all models.
arima_model_names <- c("arima_base", "arima_lagged", "arima_seasonal", "arima_seasonal_lagged")
plt <- aicc %>%
    ungroup() %>%
    rowwise() %>%
    mutate(
        mean_aicc = mean(c_across(all_of(arima_model_names)), na.rm = TRUE),
        best_model = names(.)[2 + which.min(c_across(all_of(arima_model_names)))]
    ) |>
    group_by(best_model) |>
    summarise(
        mean_aicc = mean(mean_aicc),
        n_best_model = n()
    ) |>
    mutate(mean_aicc_text = paste("Mean AICc: ", round(mean_aicc, 2))) |>
    ggplot(aes(best_model, n_best_model, fill = mean_aicc)) +
    geom_bar(stat = "identity", show.legend = FALSE) +
    geom_text(aes(label = mean_aicc_text), vjust = -0.5) +
    labs(
        title = "Best ARIMA models by AICc",
        y = "Number of best models by AICc",
        caption = glue("Models: {paste(arima_model_names, collapse = ', ')}"),
    )

render_plot(plt)
```

- The best model among arima variations is the lagged version of ARIMA.
- Prophet model has more than 60% of null models so I will remove it from the mable.
- Because NNETAR is very slow at predicting, I will remove it from the mable. Also, is not performing well in the validation set due to lack of historic data.

{{< pagebreak >}}

## RMSE

Plot best models count by minimum RMSE among models with their median.

```{r}
#| code-fold: true

plt <- rmse %>%
    ungroup() %>%
    rowwise() %>%
    mutate(
        best_model = names(.)[2 + which.min(c_across(c("arima_lagged", "stl")))],
        mean_rmse = mean(c_across(c("arima_lagged", "stl")), na.rm = TRUE),
    ) |>
    group_by(best_model) |>
    summarise(
        # The median is used instead of the mean because the median is more robust to outliers
        median_rmse = median(mean_rmse),
        n_best_model = n()
    ) |>
    mutate(median_rmse_text = paste("Median RMSE: ", round(median_rmse, 2))) |>
    ggplot(
        aes(reorder(best_model, n_best_model, decreasing = TRUE), n_best_model, fill = median_rmse)
    ) +
    geom_bar(stat = "identity", show.legend = FALSE) +
    geom_text(aes(label = median_rmse_text), vjust = -0.5) +
    labs(x = NULL, y = "Number of best models by RMSE", title = "Models performance in validation set")

render_plot(plt)
```
